---
title: "Final Project"
author: "Meet Valia"
date: "12/4/2020"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lubridate)
library(ggplot2)
library(viridis)
library(hrbrthemes)
library(dplyr)
library(zoo)
library(dplyr)
library(BAS)
library(gridExtra)
options(scipen = 999)

```

```{r}
AAPL = read.table('AAPL.csv', header = TRUE, sep = ',')

GOOG = read.table('GOOG.csv', header = TRUE, sep = ',')

FB = read.table('FB.csv', header = TRUE, sep = ',')

AMZN = read.table('AMZN.csv', header = TRUE, sep = ',')

SPY = read.table('^GSPC.csv', header = TRUE, sep = ',')
```


## Adjusted Closing Price over the Years

```{r fig.height=6, fig.width= 10, results='hide', message=FALSE, warning=FALSE}



new_APPL.df = AAPL
new_APPL.df$Date = as.Date(new_APPL.df$Date)
new_APPL.df$Year <- format(new_APPL.df$Date, "%Y")

#format(as.Date(new_df$Date, format="%d/%m/%Y"),"%Y")

adj.APPL = new_APPL.df %>%
  group_by(Year) %>%
  summarize(agg.price = sum(Adj.Close, na.rm = TRUE))



new_SPY.df = SPY
new_SPY.df$Date = as.Date(new_SPY.df$Date)
new_SPY.df$Year <- format(new_SPY.df$Date, "%Y")

#format(as.Date(new_df$Date, format="%d/%m/%Y"),"%Y")

adj.SPY= new_SPY.df %>%
  group_by(Year) %>%
  summarize(agg.price = sum(Adj.Close, na.rm = TRUE))



p1 <- ggplot(adj.APPL) +
    geom_line(aes(x=Year, y = agg.price, group = 1)) +
    scale_fill_viridis(discrete=TRUE) +
    scale_color_viridis(discrete=TRUE) +
    
    xlab("Year") +
    ylab("Adjusted Close Price") + ggtitle("Adjusted Closing Price over the Years for Apple")


p2 <- ggplot(adj.SPY) +
    geom_line(aes(x=Year, y = agg.price , group = 1)) +
    scale_fill_viridis(discrete=TRUE) +
    scale_color_viridis(discrete=TRUE) +
    
    xlab("Year") +
    ylab("Adjusted Close Price") + ggtitle("Adjusted Closing Price over the Years for GSPC")


grid.arrange(p1, p2, nrow = 1, ncol = 2)

```


## Box plot of Adjusted Closing Price of Years

```{r fig.height=6, fig.width= 10, results='hide', message=FALSE, warning=FALSE}
box.df = data.frame(AAPL$Adj.Close, new_SPY.df$Adj.Close)


colnames(box.df)[colnames(box.df) == 'AAPL.Adj.Close'] <- 'AAPL'
colnames(box.df)[colnames(box.df) == 'new_SPY.df.Adj.Close'] <- 'SPY'

box.df$AAPL = box.df$AAPL / box.df$AAPL[1]
box.df$SPY = box.df$SPY / box.df$SPY[1]

boxplot(box.df, main="Distribution of Adjusted Closing Price for Apple")

```
Since the data is normalized, and we compare the boxplots.
We can deduce from the box plot that the growth for the apple stock is higher than SPY


```{r}
stocks = SPY[, c("Date", "Adj.Close", "Low", "High")]
stocks$Date = ymd(stocks$Date)
```

```{r}
AAPL$Date = ymd(AAPL$Date)
GOOG$Date = ymd(GOOG$Date)
FB$Date = ymd(FB$Date)
AMZN$Date = ymd(AMZN$Date)
```

```{r}
names(stocks)[names(stocks) == 'Adj.Close'] = 'SPY'
names(AAPL)[names(AAPL) == 'Adj.Close'] = 'AAPL'
names(GOOG)[names(GOOG) == 'Adj.Close'] = 'GOOG'
names(FB)[names(FB) == 'Adj.Close'] = 'FB'
names(AMZN)[names(AMZN) == 'Adj.Close'] = 'AMZN'
```

```{r}
stocks = merge(x = stocks, y = AAPL[, c("Date", "AAPL")], by = "Date", all = TRUE)
stocks = merge(x = stocks, y = GOOG[, c("Date", "GOOG")], by = "Date", all = TRUE)
stocks = merge(x = stocks, y = FB[, c("Date", "FB")], by = "Date", all = TRUE)
stocks = merge(x = stocks, y = AMZN[, c("Date", "AMZN")], by = "Date", all = TRUE)
```


```{r}
names(stocks)[names(stocks) == 'High'] = 'SPY_High'
names(AAPL)[names(AAPL) == 'High'] = 'AAPL_High'
names(GOOG)[names(GOOG) == 'High'] = 'GOOG_High'
names(FB)[names(FB) == 'High'] = 'FB_High'
names(AMZN)[names(AMZN) == 'High'] = 'AMZN_High'

names(stocks)[names(stocks) == 'Low'] = 'SPY_Low'
names(AAPL)[names(AAPL) == 'Low'] = 'AAPL_Low'
names(GOOG)[names(GOOG) == 'Low'] = 'GOOG_Low'
names(FB)[names(FB) == 'Low'] = 'FB_Low'
names(AMZN)[names(AMZN) == 'Low'] = 'AMZN_Low'

stocks = merge(x = stocks, y = AAPL[, c("Date", "AAPL_High", "AAPL_Low")], by = "Date", all = TRUE)
stocks = merge(x = stocks, y = GOOG[, c("Date", "GOOG_High", "GOOG_Low")], by = "Date", all = TRUE)
stocks = merge(x = stocks, y = FB[, c("Date", "FB_High", "FB_Low")], by = "Date", all = TRUE)
stocks = merge(x = stocks, y = AMZN[, c("Date", "AMZN_High", "AMZN_Low")], by = "Date", all = TRUE)

```

```{r}
all_data = stocks

Dates = data.frame(date = all_data$Date)
```

```{r}
new_AAPL = stocks[, c("Date", "SPY", "AAPL", "AAPL_High", "AAPL_Low")]
daily.returns = c(0)
si = nrow(AAPL)
for(i in 2:si)
{
  retu = new_AAPL[i, 'AAPL'] / new_AAPL[i-1, 'AAPL'] - 1
  daily.returns = c(daily.returns, retu)
}
new_AAPL$daily.returns = daily.returns
```

## Density plot for Adjusted Closing Price for Apple

```{r fig.height=6, fig.width= 10, results='hide', message=FALSE, warning=FALSE}

p1<- ggplot(box.df, aes(x=AAPL)) +
    geom_density() +
    scale_fill_viridis(discrete=TRUE) +
    scale_color_viridis(discrete=TRUE) +
    
    xlab("Adjusted Close Price") +
    ylab("Density") + ggtitle("Density plot of Adjusted Close Price for Apple")



p2<- ggplot(new_AAPL, aes(x=daily.returns)) +
    geom_density() +
    scale_fill_viridis(discrete=TRUE) +
    scale_color_viridis(discrete=TRUE) +
    
    xlab("Daily Returns") +
    ylab("Density") + ggtitle("Density plot of Daily returns for Apple")


grid.arrange(p1, p2, nrow = 1, ncol = 2)

```




```{r}
rolling_mean <- function(arr, n=24){
  res = arr
  for(i in n:length(arr)){
    res[i] = mean(arr[(i-n):i])
  }
  res = data.frame(res)
  
  res[c(1:n-1), 1] <- NA
  res 
}

rolling_sd <- function(arr, n=24){
  res = arr
  for(i in n:length(arr)){
    res[i] = sd(arr[(i-n):i])
  }
  res = data.frame(res)
  
  res[c(1:n-1), 1] <- NA
  res 
  }

rolling_max <- function(arr, n=24){
  res = arr
  for(i in n:length(arr)){
    res[i] = max(arr[(i-n):i])
  }
  res = data.frame(res)
  
  res[c(1:n-1), 1] <- NA
  res 
  }


rolling_min <- function(arr, n=24){
  res = arr
  for(i in n:length(arr)){
    res[i] = min(arr[(i-n):i])
  }
  res = data.frame(res)
  
  res[c(1:n-1), 1] <- NA
  res 
}





AAPL_24da_mean=rolling_mean(stocks$AAPL)
AMZN_24da_mean=rolling_mean(stocks$AMZN)
AAPL_24da_add2std=AAPL_24da_mean+2*rolling_sd(stocks$AAPL)
AAPL_24da_sub2std=AAPL_24da_mean-2*rolling_sd(stocks$AAPL)
AMZN_24da_add2std=AMZN_24da_mean+2*rolling_sd(stocks$AMZN)
AMZN_24da_sub2std=AMZN_24da_mean-2*rolling_sd(stocks$AMZN)
AAPL_24da_max=rolling_max(stocks$AAPL_High)
AMZN_24da_max=rolling_max(stocks$AMZN_High)
AAPL_24da_min=rolling_min(stocks$AAPL_Low)
AMZN_24da_min=rolling_min(stocks$AMZN_Low)

```


```{r}
new_AAPL$AAPL_24da_mean = AAPL_24da_mean
new_AAPL$AAPL_24da_add2std = AAPL_24da_add2std
new_AAPL$AAPL_24da_sub2std = AAPL_24da_sub2std
new_AAPL$AAPL_24da_max = AAPL_24da_max
new_AAPL$AAPL_24da_min = AAPL_24da_min
```


```{r}
shift <- function(x, n){
  c(x[-(seq(n))], rep(NA, n))
}

Y <- shift(new_AAPL$daily.returns, 1)
Y = Y[24:2514]
```

```{r}
X = new_AAPL[24:(nrow(new_AAPL)-1), c("SPY", "AAPL_24da_mean", "AAPL_24da_add2std", "AAPL_24da_sub2std", "AAPL_24da_max", "AAPL_24da_min")]
```

```{r}
Y = data.frame(Y)
```


```{r}
shift <- function(x, n){
  c(x[-(seq(n))], rep(NA, n))
}

Y <- shift(new_AAPL$AAPL, 1)
Y = Y[24:2514]

data_df = data.frame(Y = Y, SPY = X$SPY, SMA24 = X$AAPL_24da_mean, UpperBand = X$AAPL_24da_add2std, LowerBand = X$AAPL_24da_sub2std, Max24 = X$AAPL_24da_max, Min24 = X$AAPL_24da_min)

colnames(data_df) = c("Y", "SPY", "SMA24", "UpperBand", "LowerBand", "Max24", "Min24")

```


## Co-relation between the variables


```{r fig.height=6, fig.width= 10, results='hide', message=FALSE, warning=FALSE}

variables = data_df
drops <- c("Y")
variables = variables[ , !(names(variables) %in% drops)]
pairs(variables)
```


## Split into train and test
```{r}
train_df = data_df[1:1993, ]
test_df = data_df[1994:nrow(data_df), ]
test_date = data.frame(Dates[1994:nrow(data_df), ])
```


## Bayesian Linear Regression
```{r}
stocks.bas =  bas.lm(Y ~ ., 
                   data=train_df,
                  initprobs = "marg-eplogp",
                   prior="g-prior",
                  method = "MCMC",
                   modelprior= uniform())
```


```{r}
summary(stocks.bas)

```

```{r}
coef(stocks.bas)
plot(stocks.bas, ask = F)
```


```{r}
y_test = data.frame(test_df$Y )
drops <- c("Y")
test_df = test_df[ , !(names(test_df) %in% drops)]

```


```{r}
model.selection = data.frame(model1 = 0.5012, model2 = 0.1085, model3 = 0.0827, model4 = 0.0593, model5 = 0.0515)
model.selection = data.frame(models = c("model1", "model2", "model3", "model4", "model5"), posterior.predictive.probability = c(0.5012, 0.1085, 0.0827, 0.0593, 0.0515))


ggplot(model.selection, aes(x = models, y = posterior.predictive.probability)) + geom_bar(stat = "identity") + ggtitle("Posterior probability for top 5 models")
```



## Bayesian Linear Regression Prediction

```{r}

MPM_pred <- predict(stocks.bas, newdata = test_df, estimator = "MPM", se.fit = TRUE)
ci_mpm <- confint(MPM_pred, estimator = "MPM")
opt_mpm <- which.max(MPM_pred$fit)
ci_mpm[opt_mpm, ]
#test_movie$audience_score

```


```{r}

predictions = data.frame(MPM_pred$Ybma)
predictions = cbind(predictions, test_date)



colnames(predictions)[colnames(predictions) == 'MPM_pred.Ybma'] <- 'pred'
colnames(predictions)[colnames(predictions) == 'Dates.1994.nrow.data_df....'] <- 'date'


plot.pred = cbind(predictions, y_test)
colnames(plot.pred)[colnames(plot.pred) == 'test_df.Y'] <- 'actual'


colors <- c("Actual" = "steelblue", "Predicted" = "darkred")

ggplot(plot.pred, aes(x= (date), group = 1)) + 
  geom_line(aes(y = pred, color = "Predicted")) + 
  geom_line(aes(y = actual, color="Actual"))   + xlab("Time Period") + ggtitle("Actual vs Predicted Stock Prices") + ylab("Stock Prices") + labs(colour = "Stocks") + scale_color_manual(values = colors)

```



## MSE for the Predicted Values (Bayesain Linear Regression)

```{r}
mse <- mean((predictions$pred  - y_test$test_df.Y)** 2)
mse

```


## Difference between the actual and predicted values


```{r}
difference = data.frame(predictions$pred - y_test$test_df.Y)



colnames(difference)[colnames(difference) == 'predictions.BMA.Ybma...y_test.test_df.Y'] <- 'diff'


boxplot(difference, ylab = ("Difference between the actual and predicted values"), main="Dispersion of difference between the actual and predicted values")

```


## Linear Regression


```{r}
linear = lm(Y ~ ., data = train_df)
#summary(linear)
```




## Linear  Regression Prediction

```{r fig.height=6, fig.width= 10, results='hide', message=FALSE, warning=FALSE}
lm.predict = predict(linear, newdata = test_df, se.fit = TRUE)
lm.predictions = data.frame(lm.predict$fit)

lm.predictions = cbind(lm.predictions, test_date)



colnames(lm.predictions)[colnames(lm.predictions) == 'lm.predict.fit'] <- 'linear'
colnames(lm.predictions)[colnames(lm.predictions) == 'Dates.1994.nrow.data_df....'] <- 'date'



lm.plot.pred = cbind(lm.predictions, predictions$pred)
colnames(lm.plot.pred)[colnames(lm.plot.pred) == 'predictions$pred'] <- 'bayesian'


colors <- c("Prices by Bayesian Regression" = "red", "Prices by Linear Regression" = "yellow")

ggplot(lm.plot.pred, aes(x= (date), group = 1)) + 
  geom_line(aes(y = linear, color = "Prices by Linear Regression")) + 
  geom_line(aes(y = bayesian, color="Prices by Bayesian Regression")) + xlab("Time Period") + ggtitle("Bayesian Linear Regression vs Frequentist Linear Regression Prediction") + ylab("Stock Prices") + labs(colour = "Regression Method") + scale_color_manual(values = colors)

```

## MSE for the Predicted Values (Linear Regression)

```{r}
mse <- mean((lm.predictions$linear  - y_test$test_df.Y)** 2)
mse

```
